
```{r}
setwd("C:\\Users\\panch\\Documents\\MAR 664\\PAC")
```
#Loading Data
```{r}
data = read.csv("C:\\Users\\panch\\Documents\\MAR 664\\PAC\\analysisData.csv")
scoringdata = read.csv("C:\\Users\\panch\\Documents\\MAR 664\\PAC\\scoringData.csv")
```

#Changing host_response_rate from charecter to integer
```{r}

library(readr) ; library(dplyr)

data = data%>%
  mutate(host_response_rate = parse_number(host_response_rate))

scoringdata = scoringdata%>%
  mutate(host_response_rate = parse_number(host_response_rate))

```

#Finding na cloumns in data and scoring data
```{r}

na = is.na(data)
sumofna = colSums(is.na(data))
colsumofna = colSums(is.na(data)) > 0
namesofcolswithna = colnames(data)[colSums(is.na(data)) > 0]
namesofcolswithna


na = is.na(scoringdata)
sumofna = colSums(is.na(scoringdata))
colsumofna = colSums(is.na(scoringdata)) > 0
namesofcolswithna = colnames(scoringdata)[colSums(is.na(scoringdata)) > 0]
namesofcolswithna

```
#Replacing NA values with mean of the columns in data and scoringdata
```{r}

data[is.na(data$host_listings_count),]$host_listings_count <- mean(data$host_listings_count, na.rm = T)
data[is.na(data$host_total_listings_count),]$host_total_listings_count <- mean(data$host_total_listings_count, na.rm = T)
data[is.na(data$beds),]$beds <- mean(data$beds, na.rm = T)
data[is.na(data$monthly_price),]$monthly_price <- mean(data$monthly_price, na.rm = T)
data[is.na(data$security_deposit),]$security_deposit <- mean(data$security_deposit, na.rm = T)
data[is.na(data$cleaning_fee),]$cleaning_fee <- mean(data$cleaning_fee, na.rm = T)
data[is.na(data$reviews_per_month),]$reviews_per_month <- mean(data$reviews_per_month, na.rm = T)
data[is.na(data$host_response_rate),]$host_response_rate <- mean(data$host_response_rate, na.rm = T)


scoringdata[is.na(scoringdata$host_listings_count),]$host_listings_count <- mean(scoringdata$host_listings_count, na.rm = T)
scoringdata[is.na(scoringdata$host_total_listings_count),]$host_total_listings_count <- mean(scoringdata$host_total_listings_count, na.rm = T)
scoringdata[is.na(scoringdata$beds),]$beds <- mean(scoringdata$beds, na.rm = T)
scoringdata[is.na(scoringdata$monthly_price),]$monthly_price <- mean(scoringdata$monthly_price, na.rm = T)
scoringdata[is.na(scoringdata$security_deposit),]$security_deposit <- mean(scoringdata$security_deposit, na.rm = T)
scoringdata[is.na(scoringdata$cleaning_fee),]$cleaning_fee <- mean(scoringdata$cleaning_fee, na.rm = T)
scoringdata[is.na(scoringdata$host_response_rate),]$host_response_rate <- mean(scoringdata$host_response_rate, na.rm = T)
```

 
#Converting categorical variables which are important to numeric variables of data and scoringData 
(#dont run this chunk)
```{r}

#data$instant_bookable = as.numeric(data$instant_bookable) 
#data$host_identity_verified = as.numeric(data$host_identity_verified)
#data$host_is_superhost = as.numeric(data$host_is_superhost)
#data$is_location_exact = as.numeric(data$is_location_exact)

#scoringdata$instant_bookable =as.numeric(scoringdata$instant_bookable)
#scoringdata$host_identity_verified = as.numeric(scoringdata$host_identity_verified)
#scoringdata$host_is_superhost = as.numeric(scoringdata$host_is_superhost)
#scoringdata$is_location_exact = as.numeric(scoringdata$is_location_exact)

#none are significantly correalted with price
```

#Converting categorical variables which are important to factor variables of data and scoringData
```{r}

data$room_type = as.factor(data$room_type)
data$bed_type = as.factor(data$bed_type)
data$property_type = as.factor(data$property_type)
data$cancellation_policy = as.factor(data$cancellation_policy)
data$neighbourhood_group_cleansed = as.factor(data$neighbourhood_group_cleansed)
data$host_response_time = as.factor(data$host_response_time)
data$neighbourhood = as.factor(data$neighbourhood)

scoringdata$room_type = as.factor(scoringdata$room_type)
scoringdata$bed_type = as.factor(scoringdata$bed_type)
scoringdata$property_type = as.factor(scoringdata$property_type)
scoringdata$cancellation_policy = as.factor(scoringdata$cancellation_policy)
scoringdata$neighbourhood_group_cleansed = as.factor(scoringdata$neighbourhood_group_cleansed)
scoringdata$host_response_time = as.factor(scoringdata$host_response_time)
scoringdata$neighbourhood = as.factor(scoringdata$neighbourhood)
```

#Cleaning property_type
```{r}

data$property_type[!data$property_type=='Apartment'&!data$property_type=='House'&!data$property_type=='Loft'&!data$property_type=='Condominium'&!data$property_type=='Townhouse']<-'Other'
scoringdata$property_type[!scoringdata$property_type=='Apartment'&!scoringdata$property_type=='House'&!scoringdata$property_type=='Loft'&!scoringdata$property_type=='Condominium'&!scoringdata$property_type=='Townhouse']<-'Other'

data$property_type<-as.factor(data$property_type)
scoringdata$property_type<-as.factor(scoringdata$property_type)
scoringdata$property_type<-factor(scoringdata$property_type,levels=levels(data$property_type))

levels_to_drop_property_type = c('Aparthotel',"Bed and breakfast", "Boat", "Boutique hotel" , "Bungalow", "Cabin" , "Camper/RV" , "Casa particular (Cuba)" , "Castle" , "Cave" , "Cottage" , "Dome house" , "Earth house" , "Guest suite" , "Guesthouse" , "Hostel" , "Hotel" , "Houseboat" , "Lighthouse" , "Resort" , "Serviced apartment" , "Tent" , "Tiny house","Treehouse" ,  "Villa")

#removing empty levels
data$property_type = (droplevels(data$property_type,exclude = levels_to_drop_property_type))
scoringdata$property_type = (droplevels(scoringdata$property_type,exclude = levels_to_drop_property_type))

```

#Determining if the length of different string feilds are significant or not except for amenitites and host verifications
```{r}

data$Name_Length = nchar(data$name)
data$Summary_Length = nchar(data$summary)
data$Space_Length = nchar(data$space)
data$Description_Length = nchar(data$description)
data$Neighbourhood_Length = nchar(data$neighborhood_overview)
data$Transit_Length = nchar(data$transit)
data$Access_Length = nchar(data$access)
data$Interaction_Length = nchar(data$interaction)
data$HouseRules_Length = nchar(data$house_rules)
data$HostAbout_Length = nchar(data$host_about)
data$Notes_Length = nchar(data$notes)


scoringdata$Name_Length = nchar(scoringdata$name)
scoringdata$Summary_Length = nchar(scoringdata$summary)
scoringdata$Space_Length = nchar(scoringdata$space)
scoringdata$Description_Length = nchar(scoringdata$description)
scoringdata$Neighbourhood_Length = nchar(scoringdata$neighborhood_overview)
scoringdata$Transit_Length = nchar(scoringdata$transit)
scoringdata$Access_Length = nchar(scoringdata$access)
scoringdata$Interaction_Length = nchar(scoringdata$interaction)
scoringdata$HouseRules_Length = nchar(scoringdata$house_rules)
scoringdata$HostAbout_Length = nchar(scoringdata$host_about)
scoringdata$Notes_Length = nchar(scoringdata$notes)


```

#Counting number of amenities in data and scoring data
```{r}

x = strsplit(as.character(data$amenities), ",")
y = c()
for (i in 1:length(x)){
  y[i] = length(x[[i]])
}
data$num_amenities = y


x1 = strsplit(as.character(scoringdata$amenities), ",")
y1 = c()
for (i in 1:length(x1)){
  y1[i] = length(x1[[i]])
}
scoringdata$num_amenities = y1


```

# Using amenities to create variables and finding correlation between various amentities and price
```{r}

library(stringr)
data$TV<-str_detect(data$amenities,'TV')
scoringdata$TV<-str_detect(scoringdata$amenities,'TV')
data$TV<-as.numeric(data$TV)
scoringdata$TV<-as.numeric(scoringdata$TV)



data$Cable_TV<-str_detect(data$amenities,'Cable TV')
scoringdata$Cable_TV<-str_detect(scoringdata$amenities,'Cable TV')
data$Cable_TV<-as.numeric(data$Cable_TV)
scoringdata$Cable_TV<-as.numeric(scoringdata$Cable_TV)


data$Air_conditioning<-str_detect(data$amenities,'Air conditioning')
scoringdata$Air_conditioning<-str_detect(scoringdata$amenities,'Air conditioning')
data$Air_conditioning<-as.numeric(data$Air_conditioning)
scoringdata$Air_conditioning<-as.numeric(scoringdata$Air_conditioning)

data$Paid_parking_off_premises<-str_detect(data$amenities,'Paid parking off premises')
scoringdata$Paid_parking_off_premises<-str_detect(scoringdata$amenities,'Paid parking off premises')
data$Paid_parking_off_premises<-as.numeric(data$Paid_parking_off_premises)
scoringdata$Paid_parking_off_premises<-as.numeric(scoringdata$Paid_parking_off_premises)

data$Elevator<-str_detect(data$amenities,'Elevator')
scoringdata$Elevator<-str_detect(scoringdata$amenities,'Elevator')
data$Elevator<-as.numeric(data$Elevator)
scoringdata$Elevator<-as.numeric(scoringdata$Elevator)

data$Washer<-str_detect(data$amenities,'Washer')
scoringdata$Washer<-str_detect(scoringdata$amenities,'Washer')
data$Washer<-as.numeric(data$Washer)
scoringdata$Washer<-as.numeric(scoringdata$Washer)

data$Dryer<-str_detect(data$amenities,'Dryer')
scoringdata$Dryer<-str_detect(scoringdata$amenities,'Dryer')
data$Dryer<-as.numeric(data$Dryer)
scoringdata$Dryer<-as.numeric(scoringdata$Dryer)

data$Hair_dryer<-str_detect(data$amenities,'Hair dryer')
scoringdata$Hair_dryer<-str_detect(scoringdata$amenities,'Hair dryer')
data$Hair_dryer<-as.numeric(data$Hair_dryer)
scoringdata$Hair_dryer<-as.numeric(scoringdata$Hair_dryer)

data$Iron<-str_detect(data$amenities,'Iron')
scoringdata$Iron<-str_detect(scoringdata$amenities,'Iron')
data$Iron<-as.numeric(data$Iron)
scoringdata$Iron<-as.numeric(scoringdata$Iron)

data$Laptop_friendly_workspace<-str_detect(data$amenities,'Laptop friendly workspace')
scoringdata$Laptop_friendly_workspace<-str_detect(scoringdata$amenities,'Laptop friendly workspace')
data$Laptop_friendly_workspace<-as.numeric(data$Laptop_friendly_workspace)
scoringdata$Laptop_friendly_workspace<-as.numeric(scoringdata$Laptop_friendly_workspace)

data$Dishwasher<-str_detect(data$amenities,'Dishwasher')
scoringdata$Dishwasher<-str_detect(scoringdata$amenities,'Dishwasher')
data$Dishwasher<-as.numeric(data$Dishwasher)
scoringdata$Dishwasher<-as.numeric(scoringdata$Dishwasher)

data$Bathtub<-str_detect(data$amenities,'Bathtub')
scoringdata$Bathtub<-str_detect(scoringdata$amenities,'Bathtub')
data$Bathtub<-as.numeric(data$Bathtub)
scoringdata$Bathtub<-as.numeric(scoringdata$Bathtub)

```

#Cleaning host verifications
```{r}

#removed the initial and final [] dont run this code twice 
data$host_verifications = (str_sub(data$host_verifications,2,−2))
scoringdata$host_verifications = (str_sub(scoringdata$host_verifications,2,−2))

data$host_verifications = gsub("'",'',data$host_verifications )
scoringdata$host_verifications = gsub("'",'',scoringdata$host_verifications )

a = strsplit(as.character(data$host_verifications), ",")
b = c()
for (i in 1:length(a)){
  b[i] = length(a[[i]])
}

data$num_host_verifications = b



a1 = strsplit(as.character(scoringdata$host_verifications), ",")
b1 = c()
for (i in 1:length(a1)){
  b1[i] = length(a1[[i]])
}
scoringdata$num_host_verifications = b1

#no point in exporing as cor is 0.0074
```

#Finding correlation of all numeric variables with price 
```{r}

numericVars <- which(sapply(data, is.numeric))
data_numVar <- data[,numericVars]
cor_numVar <- cor(data_numVar)
cor_sorted <- as.matrix(sort(cor_numVar[,'price'], decreasing = TRUE))

(cor_sorted) 

#all numeric variables with digits starting from second decimal are to be considered
```

#Stating the column types in the data
```{r}


NumericVars <- which(sapply(data, is.numeric)) #index vector numeric variables
FactorVars <- which(sapply(data, is.factor)) #index vector factor variables
CharacterVars <- which(sapply(data, is.character)) #index vector character variables

cat('There are', length(NumericVars), 'numeric variables, and', length(FactorVars), 'factor variables, and',length(CharacterVars),'character variables')
```
#Subsetting data and scoring data to make models
```{r}

DATA = subset(data , select = c(price,accommodates,cleaning_fee,beds,bedrooms,guests_included,bathrooms,TV,security_deposit,Dryer,Washer,Dishwasher,Cable_TV,Elevator,num_amenities,Air_conditioning,review_scores_location,Space_Length,Hair_dryer,calculated_host_listings_count_entire_homes,Iron,Paid_parking_off_premises,Laptop_friendly_workspace,Bathtub,extra_people,calculated_host_listings_count,host_listings_count,host_total_listings_count,availability_30,availability_365,Name_Length,review_scores_cleanliness,availability_60,review_scores_rating,Description_Length,availability_90,Neighbourhood_Length,Summary_Length,Notes_Length,HouseRules_Length,maximum_minimum_nights,review_scores_accuracy,minimum_nights_avg_ntm,review_scores_communication,review_scores_checkin,num_host_verifications,minimum_nights,minimum_minimum_nights,monthly_price,calculated_host_listings_count_private_rooms,property_type,room_type,bed_type,cancellation_policy,review_scores_value,HostAbout_Length,reviews_per_month,number_of_reviews,Access_Length,number_of_reviews_ltm,Interaction_Length,calculated_host_listings_count_shared_rooms))

SCORINGDATA = subset(scoringdata,select=c(accommodates,cleaning_fee,beds,bedrooms,guests_included,bathrooms,TV,security_deposit,Dryer,Washer,Dishwasher,Cable_TV,Elevator,num_amenities,Air_conditioning,review_scores_location,Space_Length,Hair_dryer,calculated_host_listings_count_entire_homes,Iron,Paid_parking_off_premises,Laptop_friendly_workspace,Bathtub,extra_people,calculated_host_listings_count,host_listings_count,host_total_listings_count,availability_30,availability_365,Name_Length,review_scores_cleanliness,availability_60,review_scores_rating,Description_Length,availability_90,Neighbourhood_Length,Summary_Length,Notes_Length,HouseRules_Length,maximum_minimum_nights,review_scores_accuracy,minimum_nights_avg_ntm,review_scores_communication,review_scores_checkin,num_host_verifications,minimum_nights,minimum_minimum_nights,monthly_price,calculated_host_listings_count_private_rooms,property_type,room_type,bed_type,cancellation_policy,review_scores_value,HostAbout_Length,reviews_per_month,number_of_reviews,Access_Length,number_of_reviews_ltm,Interaction_Length,calculated_host_listings_count_shared_rooms))
SCORINGDATA$price = NA

```

#Splitting the data into test and train
```{r}

library(caret)
set.seed(817)
split<-createDataPartition(DATA$price,p = 0.8,list=FALSE,groups = 100)
train<-DATA[split,]
test<-DATA[-split,]

```

#Best linear regression
```{r}

model_lm = lm(price ~ . , train)

pred_lm_train = predict(model_lm)
rmse_lm_train = sqrt(mean((pred_lm_train - train$price)^2)) # 71.75529

pred_lm = predict(model_lm , newdata = test)
rmse_lm_test = sqrt(mean((pred_lm - test$price)^2)) #71.37698

#pred_lm_SCORING = predict(model_lm , newdata = SCORINGDATA_1) 

rmse_lm_train ; rmse_lm_test

#(test rmse : 72.04294 , rmse after uploading : 71.58581)
```

#Best logistic regression
```{r}

model_glm = glm(price ~ ., train , family = 'poisson')

pred_glm_train = predict(model_glm, type='response')
rmse_glm_train = sqrt(mean((pred_glm_train - train$price)^2)) #70.82915

pred_glm_test = predict(model_glm , newdata = test , type = 'response')
rmse_glm_test = sqrt(mean((pred_glm_test - test$price)^2)) #71.19982

#pred_glm_SCORING = predict(model_lm , newdata = SCORINGDATA , type = 'response')

rmse_glm_train ; rmse_glm_test

#(test rmse : 70.13889 , rmse after uploading : 69.96191)
```

#Best tree selection
```{r}

library(rpart)

model_tree = rpart(price ~ . , train , method ='anova', control = rpart.control(cp = 4e-04))

pred_tree_train = predict(model_tree)
rmse_tree_train = sqrt(mean((pred_tree_train - train$price)^2)) #65.26914

pred_tree_test = predict(model_tree , newdata = test)
rmse_tree_test = sqrt(mean((pred_tree_test - test$price)^2)) #68.48148

#pred_tree_SCORING = predict(model_tree , newdata = SCORINGDATA)

rmse_tree_train ; rmse_tree_test

#(test rmse : 76.89933 , rmse after uploading : 69.96191)
```
#Adding zipcode and neighbourhood to the subset as it was causing error in linear , logistic and tree modelling
```{r}

DATA = subset(data , select =c(price,accommodates,cleaning_fee,beds,bedrooms,guests_included,bathrooms,TV,security_deposit,Dryer,Washer,Dishwasher,Cable_TV,Elevator,num_amenities,Air_conditioning,review_scores_location,Space_Length,Hair_dryer,calculated_host_listings_count_entire_homes,Iron,Paid_parking_off_premises,Laptop_friendly_workspace,Bathtub,extra_people,calculated_host_listings_count,host_listings_count,host_total_listings_count,availability_30,availability_365,Name_Length,review_scores_cleanliness,availability_60,review_scores_rating,Description_Length,availability_90,Neighbourhood_Length,Summary_Length,Notes_Length,HouseRules_Length,maximum_minimum_nights,review_scores_accuracy,minimum_nights_avg_ntm,review_scores_communication,review_scores_checkin,num_host_verifications,minimum_nights,minimum_minimum_nights,monthly_price,calculated_host_listings_count_private_rooms,neighbourhood,property_type,room_type,bed_type,cancellation_policy,review_scores_value,HostAbout_Length,reviews_per_month,number_of_reviews,Access_Length,number_of_reviews_ltm,Interaction_Length,calculated_host_listings_count_shared_rooms,zipcode))


SCORINGDATA = subset(scoringdata,select=c(accommodates,cleaning_fee,beds,bedrooms,guests_included,bathrooms,TV,security_deposit,Dryer,Washer,Dishwasher,Cable_TV,Elevator,num_amenities,Air_conditioning,review_scores_location,Space_Length,Hair_dryer,calculated_host_listings_count_entire_homes,Iron,Paid_parking_off_premises,Laptop_friendly_workspace,Bathtub,extra_people,calculated_host_listings_count,host_listings_count,host_total_listings_count,availability_30,availability_365,Name_Length,review_scores_cleanliness,availability_60,review_scores_rating,Description_Length,availability_90,Neighbourhood_Length,Summary_Length,Notes_Length,HouseRules_Length,maximum_minimum_nights,review_scores_accuracy,minimum_nights_avg_ntm,review_scores_communication,review_scores_checkin,num_host_verifications,minimum_nights,minimum_minimum_nights,monthly_price,calculated_host_listings_count_private_rooms,neighbourhood,property_type,room_type,bed_type,cancellation_policy,review_scores_value,HostAbout_Length,reviews_per_month,number_of_reviews,Access_Length,number_of_reviews_ltm,Interaction_Length,calculated_host_listings_count_shared_rooms,zipcode))
SCORINGDATA$price = NA

```


#Splitting the data into test and train after adding zipcode and neighbourhood
```{r}
 

library(caret)
set.seed(817)
split<-createDataPartition(DATA$price,p = 0.8,list=FALSE,groups = 100)
train<-DATA[split,]
test<-DATA[-split,]

```

#Best ranger model
```{r}
 
library(ranger)
set.seed(817)
ranger = ranger(price~.,
                       data = train, 
                       num.trees = 1000)

pred_ranger_train = predict(ranger, data = train, num.trees = 1000)
rmse_ranger_train = sqrt(mean((pred_ranger_train$predictions - train$price)^2))


pred_ranger_test = predict(ranger , test , num.trees = 1000)
rmse_ranger_test = sqrt(mean((pred_ranger_test$predictions - test$price)^2))

#pred_ranger_SCORING = predict(ranger , newdata = SCORINGDATA , num.trees = 1000)

rmse_ranger_train ; rmse_ranger_test

#(test rmse = 61.97403, rmse after uploading = 64.05388)
```

#Preparing data in a matrix format to feed in the XGboost model
```{r}

#install.packages("vtreat")
library(vtreat)
trt = designTreatmentsZ(dframe = train,
                        varlist = names(train)[2:64])

newvars = trt$scoreFrame[trt$scoreFrame$code%in% c('clean','lev'),'varName']

train_input = prepare(treatmentplan = trt, 
                      dframe = train,
                      varRestriction = newvars)

test_input = prepare(treatmentplan = trt, 
                     dframe = test,
                     varRestriction = newvars)

SCORING_input = prepare(treatmentplan = trt, 
                     dframe = SCORINGDATA,
                     varRestriction = newvars)
```

#Best XGboost model  
```{r}

library(xgboost)
set.seed(817)
xgboost12 <- xgboost(data = as.matrix(train_input), 
               label = train$price, 
               eta = 0.01,gamma=0,
               max_depth = 6,
               nround=6000, 
               objective="reg:squarederror",
               early_stopping_rounds=50,
               print_every_n=1000,
               subsample = 0.5,
               )


pred_xgb12_train = predict(xgboost12 , newdata = as.matrix(train_input))
rmse_xgb12_train = sqrt(mean((pred_xgb12_train - train$price)^2))

pred_xgb12_test = predict(xgboost12 , newdata = as.matrix(test_input))
rmse_xgb12_test = sqrt(mean((pred_xgb12_test - test$price)^2))

rmse_xgb12_train ; rmse_xgb12_test

#(test rmse : 57.7555, rmse after uploading = 54.51000, private rmse  = 60.28871)

```

#Forming final submission file
```{r}

pred_xgb12_SCORING = predict(xgboost12 , newdata = as.matrix(SCORING_input))

ID = scoringdata$id
PRICE = pred_xgb12_SCORING
file = data.frame(ID , PRICE)      

sub1 = write.csv(x = file , file = 'Submission_05_01_2022_4.csv' , row.names = F)
```



